Leadership across cultures and contexts: an exploratory analysis

### 2020-05-13 Response to reviewers.Rmd

**Further, and we realize there may not be sufficient power for this, but does the gender of the ethnographer predict ‘seeing’ female leaders in a population? This can (potentially) speak quantitatively to some of the qualitative discussion about limitations and biases in the approach.**

We determined the gender of almost all authors of all documents included in our study. We found that of the 884 text records with no female author or co-author, 21 (2.4%) reported female leadership. Of the 225 text records with a female author or co-author, 20 (8.9%) reported female leadership, suggesting that female ethnographers were more likely to report female leaders. After accounting for the non-independence of text records (i.e., multiple records from the same document) using a logistic mixed effects regression model, however, we found that although there was a positive effect of co-author gender on reporting female leadership (OR = 3.5), it was not statistically significant (p = 0.25). We now report this in our results section.

**Could further analyses be done to obtain more information on the nature of the variation within the data?**

In the original version of the ms we did not describe our numerous efforts at dimension reduction, including various flavors of logistic PCA, non-negative matrix factorization (NMF), and various flavors of cluster analysis, and we regressed the resulting components against our various culture-level variables. These initial efforts did not yield much convincing "stucture" in our data or results beyond what we reported. We now briefly describe these efforts so as to better contexualize the results that we do report, and we report logisticPCA and NMF results in the SI.

We now report on the dimensionality reduction that did produce stable and interpretable clusters and incorporate these "features" of leadership into our analyses and conclusions. See below.

@@

**1.	The cluster analysis is nice, but it opens up the option to combine items within both qualities and functions to create a smaller number of higher-level categories or codes that, arguably, are better measured. The issue here is that some things that don’t appear frequently might rise in frequency when they are combined with other similar items. Suppose you use a higher-level category that contains 4 codes and you code the society as ‘1’ if any of the codes appears and zero otherwise. In our view, you need not go crazy with this approach, but look for any tightly bound clusters and consider pursuing those. If you are able to do this, the presentation could be more compact and easier to digest for the reader.**

We now additionally create "feature" variables based on the clusters we did find, and investigate their relationship to our four predictor variables (i.e., our four factors).

@@

**We recommend making liberal use of supplemental materials. Some of the analyses may not yield findings of sufficient interest to warrant placement in the main text, but some readers may want to have a look at them, so put them in the supplemental and briefly mention them in the main.**

We have expanded our supplemental materials considerably.

**Returning to the WEIRD Problem: in your discussion, please circle back to the question of what research might be missing by focusing on leadership in Western or industrialized societies. Say more on how leaders in small-scale societies are the same and different—and perhaps you will be able to explain some of that variation with variables like jurisdictional hierarchy or subsistence type or mobility measures. As mentioned by R1, this likely involves returning to the GLOBE data and putting it together with your findings more directly.**

We agree we need to better situate our study within the theme of the special issue. We interpret "Beyond WEIRD" in the following way: The original WEIRD paper was a needed wake-up call for psychology. However, we now worry that many psychologists are operationalizing WEIRD as a dichotomy: WEIRD vs. non-WEIRD. This is obviously a mistake, as there is enormous variation among the non-WEIRD. Our data set focuses primarily on traditional societies, and thus speaks to variation among the non-WEIRD, a point we now emphasize. At the same time, as Edward Said famously argued, anthropologists tend to make the opposite mistake: exoticizing the "Other" relative to the "West".

We have revised our introduction section, the discussion, and our conclsuion along these lines. We hope we have struck the right balance in identifying important common themes and patterns across most of the populations in our data, while doing justice to their differences.

@@

**Status vs. Leadership: R2 asks about status vs. leadership. Given recent debates about a “leadership psychology” as distinct and separable from status psychologies (like prestige and dominance, van Vugt), it might be worth expanding a bit on what these data say about the connection between status and leadership.**

This is a good idea. We now include a short section on our "status" variable, and what distinguishes leadership records with evidence for social status from those without such evidence. We found a small set of words, such "respect", "prestige", and "honor", that were significant predictors of "high status" leaders. We found a weak effect of group structure on evidence for high status, and moderate effects of subistence strategy and continental region. We now report these results. And following this approach (an in response to other comments by reviewers) we report similar analyses for coercive authority and shamanism.

@@

**This manuscript presents a descriptive analysis of the HRAF to provide information about leadership in traditional societies. Overall, I enjoyed reading the paper and thought it was a useful resource. It provides a comprehensive analysis of leadership across a wide swath of traditional cultures. It reports on the types of functions leaders serve, the qualities possessed by leaders, and the costs and benefits of leadership experienced by both leaders and followers. The paper will likely be of interest to anyone working in the area of leadership, social hierarchy, or group behavior. I provide some comments below aimed at helping the authors increase the potential accessibility and impact of the work.**

We greatly appreciate the detailed comments and very useful suggestions.

**The paper was framed partially as providing an analysis of differences between traditional societies and WEIRD societies but, unless I missed it, there was not much analysis along these lines. Neither the background section nor the Discussion section focuses much on contrasting WEIRD and traditional societies. The paper appears to miss out on a useful opportunity to provide some commentary on this issue. Also, I realize the paper is decidedly exploratory/descriptive, but did the authors have any hypotheses about differences between WEIRD and traditional societies?**

We have now greatly revised our mansucript and more clearly explain that our exploratory analyses aim to uncover varition (or lack of variation), within "non-WEIRD" populations. We do briefly review relevant perspectives on leadership in WEIRD contexts (primarily in the introduction and to a lesser degree in our interpretations of results), but we do not aim to compare WEIRD and non-WEIRD leadership dynamics.

@@

We did not have any predictions and now more clearly state this.

**The paper suggests that one of the most important qualities and benefits of being a leader involves being of "high status." That's a little dissatisfying. What's the difference between being high status and being a leader? I realize the authors provided brief operational definitions of the two constructions, but might they simply be two ways of describing the same thing (having a high degree of influence)? Perhaps the authors could do just a bit more to differentiate them.**

We have substantially revised the mansucript and analyses regarding this point. We now include specific sections in the results, discussion, and SI which are focused on the relaitonship between leadership and high social statuts and we believe our empirical results here fit in well with much of the current evolutionary leadership literature.

@@

We also include additional details in the results and supplementary information on the corpus of ethnographic documents which provide the sample of ethnographic paragraphs.

**I don't work with ethnographic data, so perhaps this question is naïve, but: What conclusions can we draw about the actual frequency of leadership functions and traits from the frequency with which they are mentioned in the ethnographic record? There is an implicit assumption that the two are closely aligned. Has such an alignment been confirmed? Is that even testable?**

The reviewer brings up an important point. We now are more careful to frame our results as "evidence for" or "lack of evidence for" leadership traits in each culture, and to make clear that absence of evidence is not evidence of absence.

Also, this is partly our motivation for representing the percentage of cultures which provide at least one text record supporting each of the operationalized variables. Despite this low threshold for accepting supporting evidence, the ethnographic record (sampled here) does not provide support for every variable from every culture. At this culture-level we feel more confident our data provide a estimate of cross-cultural frequency, albeit an inherently noisy, "course-grained" estimation. As a further supporting point, the field of ethnology (the branch of anthropology focused on comparing characteristics of populations across the globe) is based on developing comparable and often quantitative measures from primary ethnographic information. For example, the Ethnographic Atlas, which provides hundreds of coded quantitative and qualitative variables on thousands of cultures and the Standard Cross-Cultural Sample, which provides thousands of coded quantitative and qualitative variables on 186 cultures, were developed in essentially the same way we have developed our operationalized, researcher-coded variables here. Developing and contrasting coded measures from primary ethnographic documents is not without limitations, but does represent (perhaps) the only method to systematically and empirically compare measures across a large sample of cultures/populations. At the text-record level drawing strong conclusions about the actual frequency of traits is slightly more problematic (especially the absence of relative infrequency). However, fundamentally these results provide a systematic view of the ethnographic record on these measures. Our intention is that from our results field researchers and theoreticians interested in these topics can draw inference from a more comprehensive view of human cultural diversity rather than haphazardly self-selecting and relying on only a handful of ethnographic cases studies.

@@

**The abstract could be tightened up in terms of description of particular costs and benefits to leaders and followers. And a brief justification in the abstract for why the study compares costs and benefits would be useful- e.g. understanding evolution of leader or follower strategies requires understanding the (fitness-relevant) costs and benefits to them.**

Good point. [Need to do]

**Lines 88-89 suggest that the paper's ethnographic analysis will include leadership at all levels of social organization. But does the theory and analysis reported in the paper really extend to all levels of social organization- such as child play groups or households? It seems the focus is still community-level cooperation among adults. Perhaps clarify.**

This statement in the introduction is intended to apply across disciplines focused on leadership dynamics and not specifically apply to the methods of the current study. The reviewer brings up a well-received point, however. Any work in developmental psychology on social hierarchy and leadership would not be subject to this critique. We have now rephrased this claim to better fit with our specific objectives and outcomes.

@@

We now more clearly motivate our discussion of the functions of leaders and link discussions to these sources and ideas. 
**The Wofford study (lines 214-233) and GLOBE study reminds me of the vast array of descriptive models in leadership/managerial studies- transactional vs. transformational leadership and their sub-components being one of the most popular. So what should a reader make of this literature, with respect to evolution of leadership in small-scale societies? Do bureaucracies, formal legal systems, formal education-based skills, and large inequalities in remuneration limit universality of descriptive models from leadership/managerial studies? I'd be more explicit on this in the introduction.**

We now include additional discussion of managerial frameworks of leader behavior and highlight recent approaches that are more conducive to coupling with evolutionary and anthropological/psychological approaches.

**For the discussion section: how does your own descriptive modeling of leadership in traditional societies agree or not with Wofford or other models from large-scale societies?**

We now include a paragraph in our conlcuding statement interpreting our overall findings in reference to Wofford's and GLOBE study results, i.e., attempting (albeit briefly) to tie in our (non-WEIRD) results to relevant findings from WEIRD settings.

@@

**There is little mention of gender in the Qualities of leadership section, even though this might be first attribute an alien might recognize as covarying with community-level leadership in human societies. Even in the more egalitarian societies, men often wield more overt political influence during camp or band-wide political discussions. Without getting into arguments about patriarchy and its origins, it might at least be useful to cite your own work on Chabu and work on the Tsimane (von Rueden et al. 2018) that finds similar traits predict leadership for men and women, implying its not gender per se that gives men a leadership advantage in politics.**

We now discuss sex-differences in community and political leadership, and refernce in greater deatils (albeit breif) the universal male bias in community leadership in the introuction and cite these works. We also include specific sections on female leadership and sex differences in the results and discussion.

@@

**How often are shamans also the most influential political figure (given your suggestion in line 317)? My sense from various ethnographies is that this was not the majority of cases (though often the case), whether in more egalitarian societies or chiefdoms. I could be wrong.**

We did not mean to imply shamans are often the *most* influential political leaders in this passage, only that managerial eliets (e.g., Stanish) are often associated with a special social status rooted in ideology. However, to this point we have now include a dedicated section in the results section to shaman leaders and directly address the reviewer's inquries. As we discuss, we see shamanism as a potentially important and overlooked leadership strategy.

@@

**Even though this paper is more exploratory, in the last section before the methods, can you venture any predictions about leader function-leader quality covariation? Or about how reported costs and benefits of leadership and followership may vary depending on subsistence type?**

We explicity avoided making any predictions. We understand that our review of existing evidence and theory could be used to generate a number of predictions. But ultimatley, our empirical goals are simply to explore the data structure rather than test predictions. We now more make this point more clearly.

**Many of the variables coded as qualities of leaders may also fall under leader costs and benefits, e.g. polygyny, physical health, wealthy. Were variables like these analyzed in both cases?**

The reviewer brings up an important point, i.e., the conceptual and operational overlap in many study variables, and specifically between various qualities and benifits. This is partly a consequence of our methodology, given the coding of qualities was purely driven by the ethngoraphic record whereas our coding of costs and benefits was sketched out a priori, based on evolutionary theory.

@@

Another aspect of potential overlap in the data is the quality *wealthy* and the benefit *material resources*. Here, note the operationalization of the quality *wealthy* suggests a relative comparison to followers, or a qualifying statement (we have slightly adjusted the phrasing in our table of operationalization). The benefit measures however is focused on the transactional nature of resource transfer or 'payment'.

We now discuss these disctintions in more detail in the methods and limitations sections.
<!-- NEED TO DO ABOVE? -->

**Can you control for phylogeny (e.g. linguistic affinity). See Minocher et al. 2019 in Evolution and Human Behavior. It may be that covariation you detect between qualities and leader functions is driven by shared cultural history more than direct causal relationship between qualities and leader functions.**

We appreciate this point, however as mentioned above, we did not incorporate proper phylogenetic controls. The PSF as a stratified random sample provide some safeguards against effects of non-independence, although not perfectly. We do now include significant additional analyses more systematically including and reporting continental region and subsistence strategy effects as control measures. We also now report the GLMM for each of our researcher-coded variables which includes random effects for author nested within culture and we report on possible sources of biases in the SI.

### 2020-05-13 Response to reviewers.Rmd

## Response to editor and reviewer comments

### Editor comments

**Statistical reporting and further analysis: The GLMM should be reported in full somewhere.**

The GLMM is actually 109 GLMMs (one for each coded variable). We now include a table of all GLMM coefficients in the SI.

@@

**Further, and we realize there may not be sufficient power for this, but does the gender of the ethnographer predict ‘seeing’ female leaders in a population? This can (potentially) speak quantitatively to some of the qualitative discussion about limitations and biases in the approach.**

We determined the gender of almost all authors of all documents included in our study. We found that of the 884 text records with no female author or co-author, 21 (2.4%) reported female leadership. Of the 225 text records with a female author or co-author, 20 (8.9%) reported female leadership, suggesting that female ethnographers were more likely to report female leaders. After accounting for the non-independence of text records (i.e., multiple records from the same document) using a logistic mixed effects regression model, however, we found that although there was a positive effect of co-author gender on reporting female leadership (OR = 3.5), it was not statistically significant (p = 0.25). We now report this in our results section.

**Could further analyses be done to obtain more information on the nature of the variation within the data?**

In the original version of the ms we did not describe our numerous efforts at dimension reduction, including various flavors of logistic PCA, non-negative matrix factorization (NMF), and various flavors of cluster analysis, and we regressed the resulting components against our various culture-level variables. These initial efforts did not yield much convincing "structure" in our data or results beyond what we reported. We now briefly describe these efforts so as to better contextualize the results that we do report, and we report logisticPCA and NMF results in the SI.

@@

**3.	Be clearer about how you are accounting for research effort in each society. For example, are you controlling for the total number of words or pages for each population?**

We incorporate total number of pages of ethnography for each culture. Results now reported.

**This seems very important since we want to make sure that the variation observed isn’t due to some populations getting more or less total study effort or writing effort.**

We investigated three possible sources of bias: total number of pages of ethnography, year of publication of each document, and the presence or absence of a female author/co-author for each document. We found evidence for bias in reporting of evidence for 9 of 109 variables based on year of publication only. There was no convincing evidence of bias in any variable based on the total number of pages of ethnography or ethnographer gender. Results now reported.

@@

We have expanded our supplemental materials considerably.

**Returning to the WEIRD Problem: in your discussion, please circle back to the question of what research might be missing by focusing on leadership in Western or industrialized societies. Say more on how leaders in small-scale societies are the same and different—and perhaps you will be able to explain some of that variation with variables like jurisdictional hierarchy or subsistence type or mobility measures. As mentioned by R1, this likely involves returning to the GLOBE data and putting it together with your findings more directly.**

We agree we need to better situate our study within the theme of the special issue. We interpret "Beyond WEIRD" in the following way: The original WEIRD paper was a needed wake-up call for psychology. However, we now worry that many psychologists are operationalizing WEIRD as a dichotomy: WEIRD vs. non-WEIRD. This is obviously a mistake, as there is enormous variation among the non-WEIRD. Our data set focuses primarily on traditional societies, and thus speaks to variation among the non-WEIRD, a point we now emphasize. At the same time, as Edward Said famously argued, anthropologists tend to make the opposite mistake: exoticizing the "Other" relative to the "West".

We have revised our introduction section, the discussion, and our conclusion along these lines. We hope we have struck the right balance in identifying important common themes and patterns across most of the populations in our data, while doing justice to their differences.

@@

**Status vs. Leadership: R2 asks about status vs. leadership. Given recent debates about a “leadership psychology” as distinct and separable from status psychologies (like prestige and dominance, van Vugt), it might be worth expanding a bit on what these data say about the connection between status and leadership.**

This is a good idea. We now include a short section on our "status" variable, and what distinguishes leadership records with evidence for social status from those without such evidence. We found a small set of words, such "respect", "prestige", and "honor", that were significant predictors of "high status" leaders. We found a weak effect of group structure on evidence for high status, and moderate effects of subsistence strategy and continental region. We now report these results. And following this approach (an in response to other comments by reviewers) we report similar analyses for coercive authority and shamanism.

@@

**This manuscript presents a descriptive analysis of the HRAF to provide information about leadership in traditional societies. Overall, I enjoyed reading the paper and thought it was a useful resource. It provides a comprehensive analysis of leadership across a wide swath of traditional cultures. It reports on the types of functions leaders serve, the qualities possessed by leaders, and the costs and benefits of leadership experienced by both leaders and followers. The paper will likely be of interest to anyone working in the area of leadership, social hierarchy, or group behavior. I provide some comments below aimed at helping the authors increase the potential accessibility and impact of the work.**

We greatly appreciate the detailed comments and very useful suggestions.

**The paper was framed partially as providing an analysis of differences between traditional societies and WEIRD societies but, unless I missed it, there was not much analysis along these lines. Neither the background section nor the Discussion section focuses much on contrasting WEIRD and traditional societies. The paper appears to miss out on a useful opportunity to provide some commentary on this issue. Also, I realize the paper is decidedly exploratory/descriptive, but did the authors have any hypotheses about differences between WEIRD and traditional societies?**

We have now greatly revised our manuscript and more clearly explain that our exploratory analyses aim to uncover variation (or lack of variation), within "non-WEIRD" populations. We do briefly review relevant perspectives on leadership in WEIRD contexts (primarily in the introduction and to a lesser degree in our interpretations of results), but we do not aim to compare WEIRD and non-WEIRD leadership dynamics.

@@

We did not have any predictions and now more clearly state this.

**The paper suggests that one of the most important qualities and benefits of being a leader involves being of "high status." That's a little dissatisfying. What's the difference between being high status and being a leader? I realize the authors provided brief operational definitions of the two constructions, but might they simply be two ways of describing the same thing (having a high degree of influence)? Perhaps the authors could do just a bit more to differentiate them.**

We have substantially revised the manuscript and analyses regarding this point. We now include specific sections in the results, discussion, and SI which are focused on the relationship between leadership and high social status and we believe our empirical results here fit in well with much of the current evolutionary leadership literature.

@@

**Some of the benefits of leadership also appear as costs. Greater reflection and explanation on this aspect of the data seems warranted. I understand that leaders can enjoy certain benefits (access to resources) but then if they lose their status as leader they might lose those benefits (if someone falls from grace, they no longer get a larger piece of the pie). Do the findings reflect anything more complex than that? Their discussion of benefits of costs was brief and a little disappointing. The paper might benefit from enhancing that discussion.**

We have also substantially revised our analyses of the costs and benefits of leadership, including in the results and in our interpretations. We compare the relative evidence for leader and follower benefits vs. costs in the manuscript and provide exploratory models on the total leader benefits and costs in the SI.

@@

We now discuss these limitations and points more clearly in the limitations section.

**What does it mean for someone to be a leader at the "supracommunity" level? I'm not familiar with the term, so had difficulty contextualizing the relevant findings. And how exactly do the authors define "community" versus "supracommunity"?**

In the methods section we now include a description of all levels of our *group structure* coding to clearly explain this coding scheme. We coded supracommunity leaders in cases when the "group" involves leadership beyond the community level and across multiple residential communities, such as regional political leaders, chiefs in complex chiefdoms, kings, and state-level leaders. This terminology was vaguely inspired by Murdock and Provost's (1971) Political Integration Scale which codes the political integration at the society-level as either autonomous local communities, or having one, two, or three levels of jurisdictional hierarchy "above the community"; i.e., if political authority is exclusive to a local village headman, the group would be coded as "autonomous" but if the village and headman are under authority of a chief from a separate community, the group would be coded as "one level above community", and so forth.

@@

We now more clearly motivate our discussion of the functions of leaders and link discussions to these sources and ideas. 
**The Wofford study (lines 214-233) and GLOBE study reminds me of the vast array of descriptive models in leadership/managerial studies- transactional vs. transformational leadership and their sub-components being one of the most popular. So what should a reader make of this literature, with respect to evolution of leadership in small-scale societies? Do bureaucracies, formal legal systems, formal education-based skills, and large inequalities in remuneration limit universality of descriptive models from leadership/managerial studies? I'd be more explicit on this in the introduction.**

We now include additional discussion of managerial frameworks of leader behavior and highlight recent approaches that are more conducive to coupling with evolutionary and anthropological/psychological approaches.

**For the discussion section: how does your own descriptive modeling of leadership in traditional societies agree or not with Wofford or other models from large-scale societies?**

We now include a paragraph in our concluding statement interpreting our overall findings in reference to Wofford's and GLOBE study results, i.e., attempting (albeit briefly) to tie in our (non-WEIRD) results to relevant findings from WEIRD settings.

@@

**There is little mention of gender in the Qualities of leadership section, even though this might be first attribute an alien might recognize as covarying with community-level leadership in human societies. Even in the more egalitarian societies, men often wield more overt political influence during camp or band-wide political discussions. Without getting into arguments about patriarchy and its origins, it might at least be useful to cite your own work on Chabu and work on the Tsimane (von Rueden et al. 2018) that finds similar traits predict leadership for men and women, implying its not gender per se that gives men a leadership advantage in politics.**

We now discuss sex-differences in community and political leadership, and reference in greater details (albeit brief) the universal male bias in community leadership in the introduction and cite these works. We also include specific sections on female leadership and sex differences in the results and discussion.

@@

**How often are shamans also the most influential political figure (given your suggestion in line 317)? My sense from various ethnographies is that this was not the majority of cases (though often the case), whether in more egalitarian societies or chiefdoms. I could be wrong.**

We did not mean to imply shamans are often the *most* influential political leaders in this passage, only that managerial elites (e.g., Stanish) are often associated with a special social status rooted in ideology. However, to this point we have now include a dedicated section in the results section to shaman leaders and directly address the reviewer's inquires. As we discuss, we see shamanism as a potentially important and overlooked leadership strategy.

@@

**Even though this paper is more exploratory, in the last section before the methods, can you venture any predictions about leader function-leader quality covariation? Or about how reported costs and benefits of leadership and followership may vary depending on subsistence type?**

We explicitly avoided making any predictions. We understand that our review of existing evidence and theory could be used to generate a number of predictions. But ultimately, our empirical goals are simply to explore the data structure rather than test predictions. We now more make this point more clearly.

@@

**Many of the variables coded as qualities of leaders may also fall under leader costs and benefits, e.g. polygyny, physical health, wealthy. Were variables like these analyzed in both cases?**

The reviewer brings up an important point, i.e., the conceptual and operational overlap in many study variables, and specifically between various qualities and benefits. This is partly a consequence of our methodology, given the coding of qualities was purely driven by the ethnographic record whereas our coding of costs and benefits was sketched out a priori, based on evolutionary theory.

@@

Concerning the quality of *polygynous* and the benefit of *mating*, note that the *mating* benefits includes polygyny, but also includes other mating benefits, such as having a higher quality spouse (but only one). Indeed every text recorded coded as providing evidence for the quality of *polygynous* also provides evidence for the leader *mating* benefit, however the inverse is not the case. Therefore these variables are in fact distinct.

Another aspect of potential overlap in the data is the quality *wealthy* and the benefit *material resources*. Here, note the operationalization of the quality *wealthy* suggests a relative comparison to followers, or a qualifying statement (we have slightly adjusted the phrasing in our table of operationalization). The benefit measures however is focused on the transactional nature of resource transfer or 'payment'.

We now discuss these distinctions in more detail in the methods and limitations sections.
<!-- NEED TO DO ABOVE? -->

@@

**While your analysis strategy was not designed to test particular predictions, I'd be very interested to see you report how the authority variable covaries with different leader functions in different subsistence types, which has relevance to institutionalization of leadership and evolution of social inequality.**

We now include a section in the results focused on better understanding variation with the coercive authority measure.

**Give more explanation of prestige theory and computational services theory (lines 1279-1280), including what different predictions they make.**

We now briefly (but in greater detail then our initial submission) discuss the prestige and computational services theories and interpret our empirical results in reference to these two theories. See the discussion section on "Beyond dual models."

### 2020-05-13 leadership_across_cultures_contexts.Rmd
Added leader sex by subsistence stat in text and table in SI and comment in response to reviewers.
Fixed grammer.
Re-arrange conclusions WEIRD paragraph.

The vast majority of text records pertained to male leadership (`r male_leader_pct`%) and very few pertained to female leadership (`r female_leader_pct`%), with the remaining records not sex-specific. There were female leaders in almost all types of groups, however, except military and religious. Female leaders did tend to lead at the kin group and other residential subgroups, though: `r female_residential_pct`% of female leaders led at the residential subgroup level, for instance, compared to `r male_residential_pct`% of male leaders.

@@

This study is based on the ethnographic record. Our methods are designed to capture a broad range of ethnographic descriptions from a culturally representative, systematic sample. Nonetheless, these results are conditional upon the content ethnographers chose to discuss and the way they chose to discuss it. The ethnographic record is certainly an imperfect source of information on human cultural diversity and hence a limitation of the current results. We nevertheless view the existence of the ethnographic record and the availability and functionality of the eHRAF as an enormous benefit for comparative research, given many of the populations represented in these data no longer exist or live as they did during the time of much ethnographic description. Within these ethnographic descriptions there is a general male bias, which could reflect a male bias in community leadership or the more systemic male bias in the ethnographic record [see @rosaldo_woman_1974; @mukhopadhyay_anthropological_1988]. These data are also biased towards cultural models (rather than descriptions of observed behavior) and community leadership; descriptions of kin group and residential group leadership are less frequent. The high proportion of text records discussing political group leaders could be attributable to our broad search strategy or the content of the ethnographic record of "leadership" more generally.

@@

@@

# Concluding remarks: Beyond the non-WEIRD 
<!-- NOT SURE OF TITLE -- WANT TO EMPHASIZE NON WEIRD DIVERSITY AND ALSO MENTION BACK TO WERID FINDINGS BRIEFLY -->
In recognizing the limitations of most extant psychological, sociological, and managerial data, many social scientists are now diversifying their samples to better understand *human* psychology, behavior, and culture. We caution that the WEIRD people problem in behavioral science cannot be remedied simply by a "non-WEIRD" solution. Our results, and much of anthropology and ethnography, clearly demonstrate enormous diversity within populations that are now often categorized as non-WEIRD. Furthermore, higher-order categories such as "subsistence strategy" are probably less relevant for understanding the complexity of behavioral diversity than more salient group contexts. Despite limitations, the ethnographic record presents an invaluable opportunity to explore patterns of human cultural and behavioral diversity. Evolutionary theorists and researchers looking beyond WEIRD samples should take special interest in comparative ethnographic methods.

We found leaders in the ethnographic record are portrayed as generally prosocial, are equipped with special capacities, serve to resolve conflicts and organize cooperation, and receive material, social, and reproductive benefits. These results generally provide broad cross-cultural support for several influential theories, such as the prestige-dominance model, collective action models, the service-for-prestige model, and the computational services model [for review, see @garfield_evolutionary_2019].

Our results also suggest both broad continutites and contrasts between leadership among nonindustrial, non-WEIRD populations sampled here and he WEIRD samples common in managerial studies. Our organize and strategize features of leader functions reflect the @wofford_factor_1970 *order and group achievement* and *personal interaction* factors, respectively. Also, Woffords *personal enhancement* and *dynamic and achievement* factors are similar to leadership dimensions captured by our compentencies feature. Our results, however, emphasize the prosociality of leaders in non-WEIRD settings over an emphasis on self-presevation, relative to WEIRD leaders (e.g., Woffords *security and maintenace* factor). We also identified conformity to social norms as a feature of human leadership across cultures, similar to predictions and findings drawing on Culturally Endorsed Implicit Leadership Theory by managerial scholars working with WEIRD populations [@behrendt_integrative_2017]. Our results on the functions of leaders across cultures and contexts also corresponds to the *task-oriented* vs. *relation-oriened* framework of managerial beahviors (i.e., our Managemebt vs. Prosociality features of leader functions), however, our results do not support the strong emphasis on charismastic leadership found by the GLOBE CEO study [@house2013strategic] (although charism was documented as a quality of leaders in over 20% of cultures and oratory skills in over 50%). See Figure \@ref(fig:cluser).

@@

@@

>>>>

The vast majority of text records pertained to male leadership (`r male_leader_pct`%) and very few pertained to female leadership (`r female_leader_pct`%), with the remaining records not sex-specific. There were female leaders in almost all types of groups, however, except military and religious. Female leaders did tend to lead at the kin group and other residential subgroups, though: `r female_residential_pct`% of female leaders led at the residential subgroup level, for instance, compared to `r male_residential_pct`% of male leaders. The majority of text records which exclusivley described female leaders were from horticulturlaists (`r female_hort_pct`%). See Table S\@ref(tab:sexsubtable).

@@

This study is based on the ethnographic record. Our methods are designed to capture a broad range of ethnographic descriptions from a culturally representative, systematic sample. Nonetheless, these results are conditional upon the content ethnographers chose to discuss and the way they chose to discuss it. The ethnographic record is certainly an imperfect source of information on human cultural diversity and hence a limitation of the current results. We nevertheless view the existence of the ethnographic record and the availability and functionality of the eHRAF as an enormous benefit for comparative research, given many of the populations represented in these data no longer exist or live as they did during the time of much ethnographic description. Within these ethnographic descriptions there is a general male bias, which could reflect a male bias in community leadership or the more systemic male bias in the ethnographic record [see @rosaldo_woman_1974; @mukhopadhyay_anthropological_1988]. These data are also biased towards cultural models (rather than descriptions of observed behavior) and community leadership; descriptions of kin group and residential group leadership are less frequent. The high proportion of text records discussing political group leaders could be attributable to our broad search strategy or the content of the ethnographic record of "leadership" more generally. Our sample of cultures and text records by continental region and subsistence strategy is also unbalanced.

@@

STATS limitations.

@@

# Concluding remarks: Beyond the non-WEIRD 
<!-- NOT SURE OF TITLE -- WANT TO EMPHASIZE NON WEIRD DIVERSITY AND ALSO MENTION BACK TO WERID FINDINGS BRIEFLY -->
In recognizing the limitated generalizability of most extant psychological, sociological, and managerial data, many social scientists are now diversifying their samples to better understand *human* psychology, behavior, and culture. We caution that the WEIRD people problem in behavioral science cannot be remedied simply by a "non-WEIRD" solution. Our results, and much of anthropology and ethnography, clearly demonstrate enormous diversity within populations that are now often categorized as non-WEIRD. Furthermore, higher-order explanatory variables such as "subsistence strategy" are probably less relevant for understanding the complexity of behavioral diversity than more salient group contexts. Despite limitations, the ethnographic record presents an invaluable opportunity to explore patterns of human cultural and behavioral diversity. Evolutionary theorists and researchers looking beyond WEIRD samples should take special interest in comparative ethnographic methods.

We found leaders in the ethnographic record are portrayed as generally prosocial, are equipped with special capacities, serve to resolve conflicts and organize cooperation, and receive material, social, and reproductive benefits. These results generally provide broad cross-cultural support for several influential theories, such as the dominance-prestige model, the service-for-prestige model, the computational services model, and numerous collective action models [for review, see @garfield_evolutionary_2019].

@@

Our results also suggest both broad continutites and contrasts between leadership among the nonindustrial, non-WEIRD populations sampled here and the WEIRD samples common in managerial studies. Our organize and strategize features of leader functions reflect the @wofford_factor_1970 *order and group achievement* and *personal interaction* factors, respectively. Also, Woffords *personal enhancement* and *dynamic and achievement* factors are similar to leadership dimensions captured by our compentencies feature. Our results, however, emphasize the prosociality of leaders in non-WEIRD settings over an emphasis on self-presevation, relative to WEIRD leaders (e.g., Woffords *security and maintenace* factor). We also identified conformity to social norms as a feature of human leadership across cultures, similar to predictions and findings drawing on Culturally Endorsed Implicit Leadership Theory by managerial scholars working with WEIRD populations [@behrendt_integrative_2017]. Our results on the functions of leaders across cultures and contexts also corresponds to the *task-oriented* vs. *relation-oriened* framework of managerial beahviors (i.e., our Managemebt vs. Prosociality features of leader functions), however, our results do not support the strong emphasis on charismastic leadership found by the GLOBE CEO study [@house2013strategic] (although charism was documented as a quality of leaders in over 20% of cultures and oratory skills in over 50%). See Figure \@ref(fig:cluster).

@@

```{r sexsubtable}
kable(sub_sex_tbl, 
      caption = 'Distributuon of text records by leader sex and subsistence strategy.')
```
